---
title: "Coursera Regression Models Project"
author: "Jose Antonio"
date: "28 ago 2016"
output: pdf_document
---
Motor Trend: Vehicle Fuel Consumption
Automatic x Manual tansmission
=======================================================================

** [GitHub Project Link]<https://github.com/joseantonio11/coursera-regression-model-project>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1-Executive Summary

In this project we will explore some features that affect fuel consumption in miles per gallon (MPG).
We are looking a dataset of a collection of cars (mtcars), and are interested in exploring the relationship between a set of variables and miles per gallon (MPG) that is the outcome.

In particularly we want answer two questions:
- Is an automatic or manual transmission better for MPG?
- Quantifying how different is the MPG between automatic and manual transmissions?

## 2-Describing the Data
The data to this project are extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973???74 models).
The data frame consists of 32 observations on 11 variables.

- **mpg**:	 Miles/(US) gallon
- **cyl**:	 Number of cylinders
- **disp**:	 Displacement (cu.in.)
- **hp**:	 Gross horsepower
- **drat**:	 Rear axle ratio
- **wt**:	 Weight (lb/1000)
- **qsec**:	 1/4 mile time
- **vs**:	 V/S
- **am**:	 Transmission (0 = automatic, 1 = manual)
- **gear**:	 Number of forward gears
- **carb**:	 Number of carburetors

## 3-Data Loading

```{r cars}
rm(list = ls())
library(datasets)
data(mtcars)
str(mtcars)
dim(mtcars)
```

## 4-Factor Data (wrangling)

Lets coerce the "cyl", "vs", "gear", "carb" and "am" variables into factor variables.
```{r, echo=TRUE}
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$gear <- factor(mtcars$gear)
mtcars$carb <- factor(mtcars$carb)
mtcars$am <- factor(mtcars$am)
```

For better understanding lets rename am levels to "Auto" and "Manual"
```{r}
levels(mtcars$am) <- c("Auto", "Manual")
```

## 5-Preliminary EDA

Lets look at dimension, structure and head of mtcars after the coerce as factors.
```{r}
dim(mtcars)
str(mtcars)
head(mtcars)
```

To answer our questions we are interested initially in the relation of two parameters that are Transmission (am) and Miles per Gallon (MPG). Lets plotting this relation, to evaluate it.

```{r,echo=TRUE}
## FIGURE 1 at APPENDIX
plot(mpg ~ am, data = mtcars, main = "MPG BY TRANSMISSION TYPE (AM)", xlab = "Transmission Type (AM)", ylab = "Miles Per Gallon (MPG)")
```

Visually, this plot show that manual transmissions have higher MPG's, but may have other variables that can play other role in determination of MPG, as "cyl", "disp", "hp", "wt and others.  So, could be a bias in the dataset that we are observing.
Lets plot MPG with other variables to identify correlations. 

```{r, fig.width = 10, fig.height = 10}
## FIGURE 2 AT APPENDIX
pairs(mtcars, panel = panel.smooth, main = "MTCARS PAIRS GRAPHS")
```

This graph shows that MPG has correlations with other variables than just 'am'. We need predicting MPG with relation to other variables to obtain a more accurate model. So, will be necessary to select and evaluate some models. 

## 6-Regression Analisys - Model Selection 

We may run some tests to compare the mpg with correlate variables.

### (a) Simple Linear Regression Model 
Below is the model to explain the MPG variability with the transmission type.
```{r,echo=TRUE}
fit1 <- lm(mpg ~ am, mtcars)
summary(fit1)
```
The p-value are low (0.000285) and R-Squared is 0.3385. 
Before making any conclusions on the effect of transmission type on fuel efficiency, we look at the variances between several variables in the dataset. 
Lets fitting all parameters of mtcars. 

```{r, echo=TRUE}
fitall <- lm(mpg ~ ., mtcars)
summary(fitall)
```

Reading data from Summary, there is no coefficient significant at 0.05 level and R-Squared value is not able to describe the remaining variance of the MPG variable. 
Lets use the R function STEP to do the variable selection.

### (b) STEP function

```{r}
bestFit <- step(fitall,direction="both",trace=FALSE)
summary(bestFit)
```

The Residual standard error of this model is 2.41 on 26 degrees of freedom. The Adjusted R-Squared value has increased to 0.8401 and the coefficients are significant at 0.05 significant level.

### (d) Final Model Examination

Now we fit the model "mpg ~ wt + qsec + am" as final examination model.

```{r}
lastModel <- lm(mpg ~ wt + qsec + am, data = mtcars)
summary(lastModel)$coef
```

### (d) Residual Analisys

The resulting final model examination is dependant on the transmission (am), but also weight (wt) and 1/4 mile time (qsec). All have significant p-values and the R-squared is pretty good to (0.85)

Now let's look (amongst others) at the Residuals vs Fitted 

```{r}
# FIGURE 3 AT APPENDIX
par(mfrow = c(2, 2))
plot(lastModel)
```
The Normal Q-Q plot (figure 3 APPENDIX) show that residuals are normally distributed (points close to line).
Scale-Location plot shows a constant variance due to a constant band pattern.
Residuals and Leverage shows that most of the points are contained in the 0.5 bands.

## 7 - CONCLUSION

All modes show that manual transmission will increase MPG, and this answer the first question ???Is an automatic or manual transmission better for MPG????. Based on "lastModel" using `mpg ~ wt + qsec + am` it is possible to conclude that manual transmission have more miles per gallon than automatic transmissions. With a p < 0.05 confidence cars with manual transmission have near 3 more miles per gallon than automatic transmissions.

May be if we have more observations available, with same cars model, using manual and automatic transmission, could help us better answer the second question about "Quantify the MPG difference between automatica and manual transmissions". With 32 observations it was not confidence conclude that this model fit all future observations. 

# APPENDIX

### Figure 1 : Boxplots of "mpg" vs. "am"

```{r,echo=TRUE}
## FIGURE 1
plot(mpg ~ am, data = mtcars, main = "MPG BY TRANSMISSION TYPE (AM)", xlab = "Transmission Type (AM)", ylab = "Miles Per Gallon (MPG)")
```

### Figure 2 : Pairs graph

```{r, echo=TRUE}
## FIGURE 2
pairs(mtcars, panel = panel.smooth, main = "MTCARS PAIRS GRAPHS")
```

### Figure 3 : Residual plots

```{r, echo=TRUE}
# FIGURE 3
par(mfrow = c(2, 2))
plot(lastModel)
```



